---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
########################################################################################

## Loading and preprocessing the data,loading packages

```{r read_data,echo=TRUE,results='hide'}
act_data<-read.csv("data/activity.csv")
library("dplyr")
library("timeDate")
library("lattice")
```

## What is mean total number of steps taken per day?

```{r steps_taken,echo=TRUE}
#Summarise data by date
act_data_bydate<-group_by(act_data,date)
act_data_summary<-summarise(act_data_bydate,sum(steps,na.rm=TRUE))
names(act_data_summary)[2]<-"steps"
#Create a plot
hist(act_data_summary$steps,col = "lightblue", border = "pink",main="Histogram of steps taken per day",xlab="Number of steps",ylab="Frequency")
mean_steps<-mean(act_data_summary$steps,na.rm=TRUE)
median_steps<-median(act_data_summary$steps,na.rm=TRUE)
```

The mean steps per day is `r mean_steps` and the median steps per day is `r median_steps`

## What is the average daily activity pattern?

```{r daily_pattern,echo=TRUE}
#Summarise Data by Interval
act_data_byinterval<-group_by(act_data,interval)
act_data_summary<-summarise(act_data_byinterval,mean(steps,na.rm=TRUE))
names(act_data_summary)[2]<-"steps"
#Plot the Average Daily Activity pattern
plot(act_data_summary$interval,act_data_summary$steps,type="l",main="Average daily activity pattern",xlab="Interval",ylab="Steps")
#Calculate the interval with maximum steps
max_interval<-act_data_summary$interval[act_data_summary$steps==max(act_data_summary$steps)]
```

The interval with maximum steps per day is `r max_interval`

## Inputing missing values

```{r inputing_nasteps}
head(act_data)
num_steps_NA <- length(which(is.na(act_data$steps)))
```

The number of steps with Na values is `r num_steps_NA`

```{r inputing_values}
#Group Data by date
act_data_bydate<-group_by(act_data,date)
act_data_summary<-summarise(act_data_bydate,mean(steps,na.rm=TRUE))
names(act_data_summary)[2]<-"steps"
#Replace values with NaN
act_data_summary$steps[is.nan(act_data_summary$steps)]<-0
#Create a clean dataset
act_data_clean<-act_data
for(i in 1:length(act_data_clean$steps)){
    if(is.na(act_data_clean$steps[i]))
        {act_data_clean$steps[i]<-act_data_summary$steps[act_data_summary$date %in% act_data_clean$date[i]]
             }
}
#Group by date
act_data_clean_bydate<-group_by(act_data_clean,date)
act_data_clean_summary<-summarise(act_data_clean_bydate,sum(steps,na.rm=TRUE))
names(act_data_clean_summary)[2]<-"steps"
#Plot the histogram of number of steps
hist(act_data_clean_summary$steps,col = "lightblue", border = "pink",main="Histogram of steps taken per day",xlab="Number of steps",ylab="Frequency")
mean_steps<-mean(act_data_clean_summary$steps,na.rm=TRUE)
median_steps<-median(act_data_clean_summary$steps,na.rm=TRUE)
```

The mean steps per day is `r mean_steps` and the median steps per day is `r median_steps`

**We observe that the estimates do not differ much from the first part of the Assignment because we took the mean steps in case of missing values** 

## Are there differences in activity patterns between weekdays and weekends?

```{r weekday}
#Add a column weekday toc check if it is weekday or weekend using DateTime
act_data_clean$WeekDay<-ifelse(isWeekday(as.Date(act_data$date)),"WeekDay","WeekEnd")
#Group the data by interval and WeekDay
act_data_grp<-group_by(act_data_clean,interval,WeekDay)
act_data_summary<-summarise(act_data_grp,mean(steps))
names(act_data_summary)[3]<-"steps"
#Lattice plot the Steps and interval
with(act_data_summary,xyplot(steps~interval|WeekDay,layout=c(1,2),type="l",main="Activity patterns in weekdays and weekends per interval"))
```